Notes on modeling for Kerr black holes: Basis learning, QNM frequencies, and spherical-spheroidal mixing coefficients

### 2018-06-18 docs/mvfits_paper/mvf.tex
updating doc

\par From here it is useful to note that \eqn{lin1} has a linear homogeneous matrix form. In particular, defining $\hat{U} = \{\phi_{kj}\}$, and $\vec{f} = \{f_j\}$, then $\vec{f} = \hat{U} \; \vec{\mu}$.

@@

@@

This rough algorithmic picture is spelled out in more Alg. \ref{alg:pgreedy}.

@@

  \caption{A positive (forward) greedy algorithm, \texttt{PGREEDY}. Note that a required input, $\mathcal{A}$, is a function that takes in a list of basis symbols, and outputs an estimator of fit error. In this setting, $\mathcal{A}$ is assumed to have access to peropheral information, such as the training data.}

@@

%
\subsection{Greedy Multivariate Polynomial Fitting}
%
The preceeding ...
% GMVP
\hspace{1cm}
{\scriptsize
\begin{algorithm}[H]
  %
  \caption{\gmvp{}, a degree tempered stepwise algorithm for multivariate polynomial modeling of scalar data.}
  \label{alg:pgreedy}
  %
  \begin{algorithmic}[1]
    %
    \State {\bf Input:} $ \{ \bf{x}, \bf{y} \}$
    \vskip 10pt
    \State \grey{\# Define action using matrix least squares}
    \State $\mathcal{A}$: $\epsilon = \hat{U}$
    \State Define bulk symbol space using cartesian inner-product
    \State Define list of allowed polynomial degrees
    \For { all degrees }
      \State Select all symbols with degree less than or equal to current
      \State Apply \texttt{PGREEDY} to get symbol subset and estimator val
      \If  { estimator has stalled }
        \State Break
      \EndIf
    \EndFor
    \State Apply negative greedy to get final model
    \vskip 10pt
    \State {\bf Output:} Final model
    %
  \end{algorithmic}
  %
\end{algorithm}
}

% GMVR
\hspace{1cm}
{\scriptsize
\begin{algorithm}[H]
  %
  \caption{\gmvr{}, a degree tempered stepwise algorithm for multivariate rational modeling of scalar data.}
  \label{alg:pgreedy}
  %
  \begin{algorithmic}[1]
    %
    \State {\bf Input:} $ \{ \bf{x}, \bf{y} \}$
    \vskip 10pt
    \State Define action using matrix least squares
    \State Define bulk symbol space using cartesian inner-product
    \State Define list of allowed polynomial degrees
    \For { all degrees }
      \State Select all numerator symbols with degree less than or equal to current
      \State Select all denominator symbols with degree less than or equal to current, enforcing no constant term
      \State Construct new symbol space by tagging each polynomial symbol with a boolean denoting whether it is in the numerator or denominator
      \State Apply \texttt{PGREEDY} to get symbol subset and estimator val
      \If  { estimator has stalled }
        \State Break
      \EndIf
    \EndFor
    \State Apply negative greedy to get final model
    \vskip 10pt
    \State {\bf Output:} Final model
    %
  \end{algorithmic}
  %
\end{algorithm}
}

>>>>

\par From here it is useful to note that \eqn{lin1} has a linear homogeneous matrix form.
In particular, defining $\hat{U} = \{\phi_{kj}\}$, and $\vec{f} = \{f_j\}$, then $\vec{f} = \hat{U} \; \vec{\mu}$.

@@

  \label{eq:linsol1}

@@

This rough algorithmic picture encapsulated in Alg. \ref{alg:pgreedy}.
\par Note that we will use the terms ``basis vector'' and ``model symbol'' interchangeably. However, in practice, it can be useful to represent basis vectors symbolically (e.g. with strings), and have a way of bijectively mapping between symbols and numerical basis vectors.

@@

  \caption{A positive (forward) greedy algorithm, \texttt{PGREEDY}. Note that a required input, $\mathcal{A}$, is a function that takes in a list of basis symbols, and outputs an estimator of fit error. In this setting, $\mathcal{A}$ is assumed to have access to peripheral information, such as the training data.}

@@

\subsection{Greedy Multivariate Polynomial Fitting}
The study of smooth scalar functions (e.g. $\vecf$) often centers about the Taylor series expansion.
In that instance, it is clear that any infinitely differentiable scalar function of many variables can be represented in terms of its derivatives by
  \label{eq:mvt}
  f( \vec{x} + \vec{h} ) &= e^{ \vec{x} \cdot \vec{\nabla}' } \, f(\vec{x}\,') |_{\vec{x}\,'=\vec{h}}
  &\approx \sum_{k=0}^{K} \, \frac{1}{k!} \, (\vec{x} \cdot \vec{\nabla}')^{k} \, f(\vec{x}\,') |_{\vec{x}\,'=\vec{h}}
From the first to second line of \eqn{eq:mvt}, we have used the definition of the exponential function (i.e. its series expansion form).
In the second line, the equality has been replaced by an approximation as we have limited the linear representation to $K$ terms.
This latter point is key to the perspective of \gmvp{}: given training data thought to be related to a smooth multivariate function, it may be well approximated by a truncated series expansion in the appropriate coordinate basis.
\par This truncated expansion happens to be a polynomial in at most $N$ variables.
In this setting, the uncertainty of which and how many basis terms to include makes this a problem ripe for the application of \eqn{eq:linsol1} and Alg. (\ref{alg:pgreedy}) .
\par In the current setting, the basis symbols required by Alg. (\ref{alg:pgreedy}) are the multinomial terms in \eqn{eq:mvt}.
Each term is an element of the cartesian inner-product of the power-set of each model dimension. That is $\mlam_{bulk} = \{x_0,x_1 .. x_N, ... x_0^1, x_0^2, x_0^3, ...  , (x_0^K x_1^K ... x_N^K) \}$.
Note that in practice it may be useful to encode elements of $\mlam_{bulk}$ with strings representing their constituents (e.g. $x_0 x_0 x_1 x_2 x_2 x_4$ could be represented by the string ``001224'').
\par The action, $\mathcal{A}(\mlam_{trial})$, required by Alg. (\ref{alg:pgreedy}) encompasses the evaluation of \eqn{eq:linsol1} to solve for the basis coefficients, $\mu_k$, and the calculation of the modeling error.
An explicit sktech of this is given by Alg. (\ref{alg:A_gmvp}).
  \caption{$\mathcal{A}_{\gmvp}$, the action for \gmvp. Model calculation given basis symbols, and output of model error estimate.}
  \label{alg:A_gmvp}
    \State {\bf Input:} $\mlam_{trial}$
    \State Calculate $\mu_k$ via \eqn{eq:linsol1}.
    \State Calculate the model representation error, e.g.: $\epsilon = |var(\hatU \vecmu - \vec{f})|/var(\vec{f})$, where $var$ is the variance of the respective list.
    \State {\bf Output:} $\epsilon$
\par The combination of these two ideas alone results in an algorithm prone to a deficit of stepwise methods: there is the very likely possibility that the algorithm would confuse correlated basis vectors (e.g. $x^2$ may be confused for $x^4$).
To counter this, we may incrementally increase, or \textit{temper}, the maximum allowed multinomial order.
Recall that a term such as $x_0x_1x_3^2$ has an effective order of 4.
The combination of degree tempering with the greedy approach results in the \gmvp{} algorithm as presented in Alg. (\ref{alg:gmvp})
% GMVP
  \caption{\gmvp{}, a degree tempered stepwise algorithm for multivariate polynomial modeling of scalar data.}
  \label{alg:gmvp}
    \State {\bf Input:} $ \{ \bf{x}, \bf{y}, $max\_order = 6$ \}$
    \State Define bulk symbol space using cartesian inner-product of the power set of each basis symbol
    \State Define list of allowed polynomial degrees
    \For { all degrees }
      \State Select all symbols with degree less than or equal to current
      \State Apply \texttt{PGREEDY} to get symbol subset and estimator val
      \If  { estimator has stalled }
        \State Break
    \State Apply negative greedy to get final model
    \State {\bf Output:} Final model

% GMVR
\hspace{1cm}
{\scriptsize
\begin{algorithm}[H]
  %
  \caption{\gmvr{}, a degree tempered stepwise algorithm for multivariate rational modeling of scalar data.}
  \label{alg:gmvr}
  %
  \begin{algorithmic}[1]
    %
    \State {\bf Input:} $ \{ \bf{x}, \bf{y} \}$
    \vskip 10pt
    \State Define action using matrix least squares
    \State Define bulk symbol space using cartesian inner-product
    \State Define list of allowed polynomial degrees
    \For { all degrees }
      \State Select all numerator symbols with degree less than or equal to current
      \State Select all denominator symbols with degree less than or equal to current, enforcing no constant term
      \State Construct new symbol space by tagging each polynomial symbol with a boolean denoting whether it is in the numerator or denominator
      \State Apply \texttt{PGREEDY} to get symbol subset and estimator val
      \If  { estimator has stalled }
        \State Break
      \EndIf
    \EndFor
    \State Apply negative greedy to get final model
    \vskip 10pt
    \State {\bf Output:} Final model
    %
  \end{algorithmic}
  %
\end{algorithm}
}
